A meta-analysis of powdery mildew yield loss mitigation strategies in Australian mungbean

### 2020-07-14 07_Manuscript.Rmd
writing corrections, loading data and plots for figures tables and hard-coded results

# Data
PM_Mdat <- read.csv(here("cache/PM_MB_updated.csv"))
```

@@

## Materials and Methods
<!-- This is where I am working right now (Paul) :) :) :) -->
The data for this study were obtained from through personal correspondence with colleagues and collaborating institutions.
The majority of the trials were undertaken by Professor Malcolm Riley and Dr Sue Thompson with the Queensland Department of Agriculture, Fisheries and Forestry (DAF). 
Trials undertaken in the 2013 season were conducted by the Northern Growers Alliance (NGA) [@goolhi2013; @premer2013; @Millmerran2013; @Marysmount2013].
<!-- should check, Sue and Mal moved to USQ in 2015 I think. Therefore the trials probably started with USQ earlier --> The most recent trials starting from 2015 were undertaken by the University of Southern Queensland (USQ) with the assistance of Queensland DAF from 2015 to 2019. 
Trial data were in different formats with varying levels of information.
Some trial data reported means of each treatment, with all but one study reporting some form of variance along with the mean.

@@

@@

@@

<!-- Write about first sign of disease in March -->
Overall the meta-analysis indicated that, with the exception of `Early` single sprays all the the spray schedules were effective at protecting mungbean yields, relative to the no spray controls (Figure 1). 
Treatments which incorporating multiple fungicide applications, `RecommendedPlus` and `LatePlus` produced the highest mean yield protection estimates at 196.4 (SE = 48.7&nbsp;kg&nbsp;/&nbsp;ha) and 219.9&nbsp;kg&nbsp;/&nbsp;ha (SE = 47.4&nbsp;kg&nbsp;/&nbsp;ha) respectively (Table 3). 
Both `RecommendedPlus` and `LatePlus` were also significantly higher than single applications starting `Early` (P = 0.008 and P = 0.003 respectively) and `Late`(P = 0.05 and P = 0.03 respectively). 
However, comparing `LatePlus` and `Early` should done with caution given these treatments did not occur within the same trial for any of the 15 trials included in the meta-analysis. 
Within the single fungicide application treatments `Recommended` provided the highest yield protection, and was estimated at saving an average of 159.9 kg / ha (SE = 30.5 kg / ha) relative to the no spray control (P < 0.0001). 
`Recommended` treatments were was also significantly higher than `Early` treatments by 122.1 kg / ha (SE = 42.1 kg / ha)(P = 0.004). 
Single `Late` applications were effective at significantly mitigating the effect of powdery mildew on yield by a mean of 122 kg / ha (SE = 39.2 kg / ha)(P = 0.0019). 
However, this estimate was not significantly different to single `Early` or `Recommended` treatments (P = 0.119; P = 0.278 respectively).

@@

@@

## Discussion
The current advice provided to growers for when to commence fungicide applications for control of powdery mildew is at the first sign of the disease [@SueThompson2016]. This meta-analysis confirms this advice by showing that applying the first fungicide application, within three days of first sign (Recommended) and up to 13 days for as was seen in the `Late` treatments. In addition spray schedules with follow-up fungicide applications made at least two weeks after the first application were on average more effective at reducing yield loss due to powdery mildew, which again follows the advice given by Thompson et. al [@SueThompson2016]. Fungicide applications which were made prior to any sign of the disease we not effective at saving yield.

>>>>

# Data
PM_Mdat <- read.csv(here("cache/PM_MB_updated.csv"))
load(file = here("cache/Meta-analysisData.Rdata"))
load(file = here("cache/PreliminaryData.Rdata"))
```

@@

## Materials and Methods
The data for this study were obtained from through personal correspondence with colleagues and collaborating institutions.
The majority of the trials were undertaken by Professor Malcolm Riley and Dr Sue Thompson with the Queensland Department of Agriculture, Fisheries and Forestry (DAF). 
Trials undertaken in the 2013 season were conducted by the Northern Growers Alliance (NGA) [@goolhi2013; @premer2013; @Millmerran2013; @Marysmount2013].
<!-- should check, Sue and Mal moved to USQ in 2015 I think. Therefore the trials probably started with USQ earlier --> The most recent trials starting from 2015 were undertaken by the University of Southern Queensland (USQ) with the assistance of Queensland DAF from 2015 to 2019. 
Trial data were in different formats with varying levels of information.
Some trial data reported means of each treatment, with all but one study reporting some form of variance along with the mean.

@@

<!-- Passage here about trial sowing times and first sign of disease (Figure 1)  PM -->

@@

```{r Figure2_treatForest}
YieldContrasts
```

@@

<!-- Write about first sign of disease in March -->
Overall the meta-analysis indicated that, with the exception of `Early` single sprays all the the spray schedules were effective at protecting mungbean yields, relative to the no spray controls (Figure 2). 
Treatments which incorporating multiple fungicide applications, `RecommendedPlus` and `LatePlus` produced the highest mean yield protection estimates at 196.4 (SE = 48.7&nbsp;kg&nbsp;/&nbsp;ha) and 219.9&nbsp;kg&nbsp;/&nbsp;ha (SE = 47.4&nbsp;kg&nbsp;/&nbsp;ha) respectively (Table 3). 
Both `RecommendedPlus` and `LatePlus` were also significantly higher than single applications starting `Early` (P = 0.008 and P = 0.003 respectively) and `Late`(P = 0.05 and P = 0.03 respectively). 
However, comparing `LatePlus` and `Early` should done with caution given these treatments did not occur within the same trial for any of the 15 trials included in the meta-analysis. 
Within the single fungicide application treatments `Recommended` provided the highest yield protection, and was estimated at saving an average of 159.9 kg / ha (SE = 30.5 kg / ha) relative to the no spray control (P < 0.0001). 
`Recommended` treatments were was also significantly higher than `Early` treatments by 122.1 kg / ha (SE = 42.1 kg / ha)(P = 0.004). 
Single `Late` applications were effective at significantly mitigating the effect of powdery mildew on yield by a mean of 122 kg / ha (SE = 39.2 kg / ha)(P = 0.0019). 
However, this estimate was not significantly different to single `Early` or `Recommended` treatments (P = 0.119; P = 0.278 respectively).

@@

```{r Figure1}
SeasonPlot
```

@@

## Discussion
The current advice provided to growers for when to commence fungicide applications for control of powdery mildew, is spray at the first sign of the disease and if necessary a follow-up spray two weeks later [@SueThompson2016; @Sparks2017]. 
This meta-analysis confirms that advice, by showing that applying the first fungicide application, within three days of first sign will significantly mitigate yield loss from powdery mildew ($Recommended$, P = `r signif(PM_mv$pval[3],2)`). 
However in addition to this the analysis also showed spray schedules which commenced 'late', up to 13 days after first sign of the disease was still effective at mitigating yield loss ($Late$, P =  `r signif(PM_mv$pval[5],2)`). 
Additionally, spray schedules with follow-up fungicide applications made at least two weeks after the first application were on average more effective at reducing yield loss due to powdery mildew, which again follows the advice given by Thompson et. al [@SueThompson2016].
This meta-analysis adds certainty to the advice which is currently given to growers, especially around any uncertainty that Early sprays might be effective at suppressing the disease. 
We show here that Early fungicide applications, which were made prior to any sign of the disease, were not effective at mitigating yield loss.

The results also provide a effective spray window for mitigating yield losses, given that spray schedules commenced `Late` were still significantly effective at mitigating yield loss (P = `signif(PM_mv$pval[5],2)`) at least up to 13 days after the first sign of disease. 
Growers typically don't have the time to inspect their crop regularly for the first sign of powdery mildew. 
Also weather conditions and logistical issues might prevent them from getting machinery in the paddock to treat the disease immediately after it is spotted. 
These issues and perhaps other on farm duties might delay a timely application, which might permit the disease time to establish. 
Knowing a delay of up to two weeks will not lead to significant yield losses can allow growers to incorporate other integrated pest management options, such as insecticide, with the fungicide spray saving on spray application costs. 
However care should be taken to ensure if a second follow-up fungicide application is necessary, sufficient time should be permitted to schedule the spray and avoid the end of season withholding periods on fungicides.

It is important to note, the studies included in this meta-analysis were intentionally sown late to ensure infection during the experiment. 
While growers may have to choose sowing mungbeans late, if the crop is sown too late there is a risk of it being frosted-out.
Regardless in most cases growers will likely have sown earlier than the trials used here reducing the risk of powdery mildew, and also reducing the requirements for fungicide applications.
Despite the confidence in the result of this model discretion is still required for if fungicide applications are necessary given the crop growth stage and forecast weather conditions.
<!-- Insert Powdery mildew plug here-->
<!-- This is where I am working right now (Paul) :) :) :) -->

These results suggest that when weather conditions cool and become suitable for powdery mildew growers should scout their every week to two weeks
